Reasoning about Connectors in Coq

### 2017-05-29 FACS_Extension/paper_revised0508.tex
code configuration, algorithm and some details revised

%\maketitle
\begin{abstract}
Reo is a channel-based exogenous coordination model in which complex coordinators, called connectors, are compositionally
built out of simpler ones. In this paper, we present a new approach to model connectors in Coq which is a proof assistant based on higher-order logic and $\lambda$-calculus.
The model reflects the original structure of connectors simply and clearly. In our framework, basic connectors (channels) are interpreted as axioms and composition operations are specified as inference rules. Furthermore, connectors are interpreted as logical predicates which describe the relation
between inputs and outputs. With such definitions provided, connector properties, as well as equivalence and refinement relations between different connectors, can be naturally formalized as \emph{goals} in Coq and easily proved using pre-defined \emph{tactics}. Particularly, to fulfil the proof of refinement relation, Z3, which can be used as a component in the context of other tools that require solving logical formulas, works as assistance.
%\paragraph*{Keywords:}

@@

In this paper, we aim to provide an approach to formally modeling and reasoning about connectors using Coq.
The basic idea of our
approach is to model the behavior of a connector by representing it as a logical predicate which describes the relation among
the timed data streams on the input and output nodes, and to reason about connectors' properties, as
well as the equivalence and refinement relations
between connectors, by using proof principles and tactics in Coq.
Compared with existing approaches for verifying connectors' properties \cite{BBK+10,KB09,KKV12}, using Coq is especially helpful when
we take infinite behavior into consideration. The coinductive proof principle makes it possible to prove connectors' properties easily while it
is difficult (sometimes impossible) for other approaches (like model checking) because of the huge (or maybe infinite) number of states. The main drawback of failing to check the complementary refinement relation in Coq is made up by Z3.
While Coq provides a platform to prove properties with a series of tactics and strategies to be considered, Z3 can be used as a component in the context of other tools that require solving logical formulas which is in consistent with the definition of Reo connectors.

@@

The paper is organized as follows: After this general introduction, we briefly summarize Reo and Coq
in Section \ref{sec:pre}. Section \ref{sec:basicdef} shows the notion of timed data streams and some
pre-defined auxiliary functions and predicates. Section \ref{sec:channelandoperator} presents the
formal modeling of basic channels and operators, as well as complex connectors. Section
\ref{sec:verification} shows how to reason about connector properties and equivalence (or
refinement) relations in our framework. Section \ref{sec:refinement} concentrates on the proof of refinement relation between two connectors.
In Section \ref{sec:conclusion}, we concludes with some further research directions. Full source codes can be found at \cite{reo2coqfile} for further reference.

@@

In this section, we provide a brief introduction to the coordination language Reo and proof assistant tools Coq and Z3.

@@

%\subsection{Timed data streams}
The behavior of a connector can be formalized by means of data-flows at its sink and source nodes which are essentially infinite sequences. With the help of the stream library in Coq, such infinite data-flows can be defined as \emph{timed data streams}:
\begin{verbatim}
    Definition Time := R.
    Definition Data := nat.
    (*Inductive Data : Set :=
        |Natdata : nat-> Data
        |Empty : Data.*)
    Definition TD := Time * Data.
    Variable Input : Stream TD.
    Variable Output : Stream TD.
\end{verbatim}
%
In our framework, time is represented by real numbers. Benefit from the completeness of real number system, we can express and carry out the effective operation of a quantity at any precision request.
The continuity of the set of real numbers is sufficiently enough for our modeling approach. Also the
continuous time model is more appropriate since it is very expressive and closer to the nature of time in the real world. Thus, the time sequence consists of increasing and diverging time moments. For simplicity, here we take the natural numbers as the definition of data, which can be easily expanded according to different application domains. The Cartesian product of time and data defines a TD object.
We use the stream module in Coq to produce streams of TD objects.

@@

The simplest form of a synchronous channel is denoted by the Sync channel type. For a channel of the Sync type, a read operation on
its sink end succeeds only if there is a write operation pending on its source end. Thus, the time and data of a stream flowing
into the channel are exactly the same as the stream that flows out of the channel
\footnote{If we use $\alpha,\beta$ to denote the data streams that flow through the channel ends of a channel and $a,b$ to denote the time stream corresponding to the data streams, i.e., the $i$-th element $a(i)$ in $a$ denotes exactly the time moment of the occurrence of $\alpha(i)$, then we can easily obtain the specifications for different channels, as discussed in \cite{Sun12,SAA+12}. For example, a synchronous channel can be expressed as $\alpha=\beta \wedge \emph{a}=\emph{b}$.}.
The Sync channel can be defined as follows in the Coq system:
\begin{verbatim}
    Definition Sync (Input Output:Stream TD) : Prop :=
    Teq Input Output /\ Deq Input Output.
\end{verbatim}

The channel of type SyncDrain is a synchronous channel that allows pairs of write operations pending on its two ends to succeed simultaneously. All written data items are lost. Thus, the SyncDrain channel is used for synchronising two timed data streams on its two source ends. This channel type is an important basic synchronization building block for the construction of more complex connectors. The SyncDrain channel can be defined as follows:
\begin{verbatim}
    Definition SyncDrain (Input Output:Stream TD) : Prop :=
    Teq Input Output.
\end{verbatim}

The channel types FIFO and FIFO$n$ where $n$ is an integer greater than $0$ represent the
typical unbounded and bounded asynchronous FIFO channels. A write to a FIFO channel always succeeds, and a write to a FIFO$n$ channel succeeds only if the
number of data items in its buffer is less than its bounded capacity $n$. A read or take from a FIFO or FIFO$n$ channel suspends until the first data item
in the channel buffer can be obtained and then the operation succeeds.
%We use the $\overrightarrow{\emph{a}}$ to denote the tail of  a sequence \emph{a}.
For simplicity, we take the FIFO1 channel as an example. This channel type requires that the time when it consumes a data item through its source end is earlier than the time when the data item is delivered through its sink end. Besides, as the buffer has the capacity 1, time of the next data item that flows in should be later than the time when the data in the buffer is delivered. We use intersection of predicates in its definition as follows:
\begin{verbatim}
    Definition FIFO1(Input Output:Stream TD) : Prop :=
    Tle Input Output /\ Tle Output (tl Input)
    /\ Deq Input Output.
\end{verbatim}

For a FIFO1 channel whose buffer already contains a data element \emph{e}, the communication can be initiated only if the data element \emph{e} can be taken via the sink end. In this case, the data stream that flows out of the channel should get an extra element \emph{e} settled at the beginning of the stream. And time of the stream that flows into the channel should be earlier than time of the tail of the stream that flows out. But as the buffer contains the data element \emph{e}, new data can be written into the channel only after the element \emph{e} has been taken. Therefore, time of the stream that flows out is earlier than time of the stream that flows in. The channel can be represented as the intersection of several predicates as follows:
\begin{verbatim}
    Definition FIFO1e(Input Output:Stream TD)(e:Data) : Prop :=
    Tgt Input Output /\ Tle Input (tl Output)
    /\ PrR (hd Output) = e  /\ Deq Input (tl Output).
\end{verbatim}

@@

A LossySync channel behaves the same as a Sync channel, except that a write operation on its source always succeeds immediately. If a compatible read or take operation is already pending on the sink of a LossySync channel, the written data item is transferred to the pending operation and both succeed. Otherwise, the write operation succeeds and the data item is lost. The LossySync channel can be defined as follows:
\begin{verbatim}
   Parameter LossySync: Stream TD -> Stream TD -> Prop.
   Axiom LossySync_coind:
     forall Input Output: Stream TD,
     LossySync Input Output ->
       (
        (hd Output = hd Input  /\ LossySync (tl Input)(tl Output))
        \/
        LossySync(tl Input) Output
       ).
\end{verbatim}

AsyncDrain is analogous to SyncDrain except that it guarantees that the pairs of write operations on the two channel ends never succeed simultaneously. Similarly it only has requirements on the time of the two streams on its opposite ends, but it requires that the times of the two streams are always different. The AsyncDrain channel can be defined as follows:
\begin{verbatim}
    Parameter AsyncDrain: Stream TD -> Stream TD -> Prop.
    Axiom AsyncDrain_coind:
      forall Input1 Input2: Stream TD,
      AsyncDrain Input1 Input2 ->
        (~ PrL(hd Input1)  =  PrL (hd Input2) )
        /\
        ( (
          (PrL(hd Input1) < PrL (hd Input2)) /\
          AsyncDrain (tl Input1) Input2
          ) /\
          (
          (PrL(hd Input1) > PrL (hd Input2))  /\
          AsyncDrain Input1 (tl Input2)
          )
        ).
\end{verbatim}

@@

The merge operator is more complicated. We consider merging two
channels \emph{AB} and \emph{CD}. When the merge operator acts on
these two channels, it leads to a choice of taking from the common
node that delivers a data item out of \emph{AB} or \emph{CD}.
%If both channels have data items available, the choice is
%non-deterministic\footnote{Choices are made by external context, so
%the definition here remains deterministic.}.
Similar to the definition of basic channels, we define merge as the intersection of two predicates and use recursive definition here:
%    Definition xor(a b: Prop) :=  (a \/ b) /\ ~(a /\ b).
\begin{lstlisting}
    Parameter merge:
    Stream TD -> Stream TD ->Stream TD -> Prop.
    Axiom merge_coind:
      forall s1 s2 s3:Stream TD,
      merge s1 s2 s3-> (
        ~ (PrL(hd s1) = PrL(hd s2)) /\
        (
          (PrL(hd s1) < PrL(hd s2)) ->
          ((hd s3 = hd s1)  /\ merge (tl s1) s2 (tl s3))
        ) /\ (
          (PrL(hd s1) > PrL(hd s2)) ->
          ((hd s3 = hd s2)  /\ merge s1 (tl s2) (tl s3))
        )
      ).
\end{lstlisting}

@@

In Coq, the theorem can be proved as follows:
\begin{verbatim}
    Theorem test1: forall A B C,
      Sync A B /\ FIFO1 B C -> Tle A C.
    Proof.
        intros. destruct H. destruct H0.
        intro n. rewrite H. apply H0.
    Qed.
\end{verbatim}

@@

We first introduce some lemmas to facilitate the proof.
\begin{verbatim}
    Lemma transfer_eq : forall s1 s2 s3 : Stream TD,
    ((Teq s1 s2) /\ (Teq s2 s3)) -> (Teq s1 s3).
    Lemma transfer_eqtl : forall s1 s2 : Stream TD,
    (Teq s1 s2) -> (Teq tl s1)) (tl s2)).
    Lemma transfer_leeq : forall s1 s2 s3 : Stream TD,
    ((Tle s1 s2) /\ (Teq s2 s3)) -> (Tle s1 s3).
    Lemma transfer_hdle : forall s1 s2 : Stream TD,
    (Tle s2 s1) -> (PrL (hd s1) > PrL (hd s2)).
\end{verbatim}

@@

\ionode{(io1)}{(1,1.2)}{node[left]{A}}
\ionode{(io2)}{(1,0)}{node[left]{B}}
\ionode{(io3)}{(4.5,1.2)}{node[right]{C}}
%\ionode{(io4)}{(5,0)}{node[right]{C2}}
\syncdrain{(io1)}{(io2)}{}
\fifoe{(io1)}{(io3)}{}
\sync{(io2)}{(io3)}{}
\draw (2.7, -0.6) node[draw=none] {(b) After merge};
\end{tikzpicture}
\caption{Alternator}
\label{fig:alternator}
\end{figure}

@@

\begin{verbatim}
    Section Alt.
    Hypothesis D1: SyncDrain A B.
    Hypothesis D2: FIFO1 A C1.
    Hypothesis D3: Sync B C2.
    Theorem subtest:
       (Tle C2 C1)  /\  (Tle C1 (tl C2)).
\end{verbatim}

@@

An additional hypothesis is needed for the proof of alternator which merges \emph{C1} and \emph{C2} into a common node \emph{C}. Based on the three
hypotheses for channels and the additional hypothesis, the theorem of alternator is presented as the following proposition which needs to be proved:
\begin{verbatim}
    Hypothesis D4: merge C1 C2 C.
    Theorem test:
    hd(C) = hd(C2) /\ merge C1 (tl C2) (tl C).
    Proof.
      destruct subtest.  (* ... *)
\end{verbatim}
Here we only present the first step which shows how a proven theorem can be applied in another proof and omit the full details because of the page limitation.
And it greatly simplifies the process of proving the property of alternator.
\end{example}

@@

\begin{verbatim}
    Theorem refinement : forall A C D,
    (exists B, (FIFO1 A B) /\ (Sync B C) /\ (Sync B D)) ->
    (exists E, (Sync A E) /\ (FIFO1 E C) /\ (FIFO1 E D)).
\end{verbatim}

@@

Equivalence relationship between the two connectors can be formalized as:
\begin{verbatim}
    Theorem equivalence: forall A B C,
      (exists E F G,
        (Sync A E) /\ (FIFO1 E F) /\ (Sync F B) /\
        (FIFO1 E G) /\ (Sync G C) /\ (SyncDrain F G)
      ) <->
      (exists D,
        (FIFO1 A D) /\ (Sync D B) /\ (Sync D C)
      ).
\end{verbatim}
The proof of this theorem has two steps. Firstly, we prove that the new connector \emph{R} is a refinement of connector \emph{Q}.
We hope to find an appropriate \emph{$D$} that satisfies
\[
FIFO1(A,D)\wedge Sync(D,B) \wedge Sync(D,C)
\]
Similar to Example \ref{ex:refine}, we first assert $D=F$, which leads to
\[
FIFO1(A,F) \land Sync(F,B) \wedge Sync(F,C)
\]
From Lemma \ref{lemma:eq}, we have $Sync(A,E)$, or $A=E$. Therefore, \emph{FIFO1(E,F)} can be replaced by \emph{FIFO1(A,F)}. By adopting \emph{FIFO1(E,F)} and \emph{FIFO1(E,G)}, we can prove that the data sequences at \emph{F} and \emph{G} are equal. Similarly, data sequences at $C$, $G$ and $F$ are also equal, wrt. $Sync(G,C)$.

@@

By means of function \emph{Conjunction}, basic channels are defined concisely and neatly. But the biggest difference is that what we feed Z3 solver need to be constraints such that Z3 solver can make decisions and yield results about refinement relations between connectors. For example, the specific behavior constraints for \emph{Sync} channel are classified into two types: data constraints and time constraints. The description and the definition of \emph{Sync} in Coq have already clearly expressed the behavior pattern it needs to follow: each item in data and time streams of output specified as ``node[1]'' is equivalent to the data and time streams of input specified as ``node[0]'', which are exactly data and time constraints. As they are both requirements we need to satisfy no matter data or time is related, each constraint in the list will all be merged finally.
\lstset{language=Python}
\begin{lstlisting}[frame=single]
def Sync(nodes, bound):
   assert len(nodes) == 2
   constraints = []
   for i in range(bound):
     constraints += [ nodes[0]['data'][i] == nodes[1]['data'][i] ]
     constraints += [ nodes[0]['time'][i] == nodes[1]['time'][i] ]
   return Merge(constraints)
\end{lstlisting}

@@

c2 = Connector()
c2.connect('Fifo1', 'A', 'N')
c2.connect('Sync', 'N', 'B')
c2.connect('Sync', 'N', 'C')
\end{lstlisting}
Notice that the \emph{replicate} operator is applied twice in the construction process. Assume the bound is ten and then carry out the function \emph{c2.isRefinementOf(c1, 10)}. Finally, we can get the result \emph{True} which means that \emph{Connector Q} in Fig.\ref{refine} is indeed a refinement of \emph{Connector P}. We can further extend the bound and the time complexity increases linearly as no recursively defined channels are involved.
\end{example}
\begin{example}
This example checks the equivalence relation between connectors in Fig.\ref{refine1}. Construct the two connectors first similar to the example \ref{ex:equivalence1} and invoke the function \emph{isRefinementOf} in two directions. Then we can get the equivalence result.
When giving the bound ten, the results of two directions are both \emph{True}. Hence, we can conclude that the two connectors are in equivalence relation.
\end{example}

>>>>

%\maketitle
\begin{abstract}
Reo is a channel-based exogenous coordination model in which complex coordinators, called connectors, are compositionally
built out of simpler ones. In this paper, we present a new approach to model connectors in Coq which is a proof assistant based on higher-order logic and $\lambda$-calculus.
The model reflects the original structure of connectors simply and clearly. In our framework, basic connectors (channels) are interpreted as axioms and composition operations are specified as inference rules. Furthermore, connectors are interpreted as logical predicates which describe the relation
between inputs and outputs. With such definitions provided, connector properties, as well as equivalence and refinement relations between different connectors, can be naturally formalized as \emph{goals} in Coq and easily proved using pre-defined \emph{tactics}. Particularly, to fulfil the proof of refinement relation, Z3 is used as an assistant tool for solving logical formulas.
%\paragraph*{Keywords:}

@@

In this paper, we aim to provide an approach to formally modeling and reasoning about connectors using Coq.
The basic idea of our
approach is to model the behavior of a connector by representing it as a logical predicate which describes the relation among
the timed data streams on the input and output nodes, and to reason about connectors' properties, as
well as the equivalence and refinement relations
between connectors, by using proof principles and tactics in Coq.
Compared with existing approaches for verifying connectors' properties \cite{BBK+10,KB09,KKV12}, using Coq is especially helpful when
we take infinite behavior into consideration. The coinductive proof principle makes it possible to prove connectors' properties easily while it
is difficult (sometimes impossible) for other approaches (like model checking) because of the huge (or maybe infinite) number of states. The main drawback of failing to check the complementary refinement relation in Coq is made up by Z3.
While Coq provides a platform to prove properties with a series of tactics and strategies to be considered, Z3 can be used as a component in the context of other tools that require solving logical formulas. The logical formulas involved in the model of Reo should reflect the constraints of connectors and be consistent with their behavior.

@@

The paper is organized as follows: After this general introduction, we briefly summarize Reo, Coq and Z3
in Section \ref{sec:pre}. Section \ref{sec:basicdef} shows the notion of timed data streams and some
pre-defined auxiliary functions and predicates. Section \ref{sec:channelandoperator} presents the
formal modeling of basic channels and operators, as well as complex connectors. Section
\ref{sec:verification} shows how to reason about connector properties and equivalence (or
refinement) relations in our framework. Section \ref{sec:refinement} concentrates on the proof of refinement relation between two connectors.
In Section \ref{sec:conclusion}, we concludes with some further research directions. Full source codes can be found at \cite{reo2coqfile} for further reference.

@@

In this section, we provide a brief introduction to the coordination language Reo, proof assistant tools Coq and Z3.

@@

%\subsection{Timed data streams}
The behavior of a connector can be formalized by means of data-flows at its sink and source nodes which are essentially infinite sequences. With the help of the stream library in Coq, such infinite data-flows can be defined as \emph{timed data streams}:
\begin{lstlisting}[language=coq]
    Definition Time := R.
    Definition Data := nat.
    (*Inductive Data : Set :=
        |Natdata : nat-> Data
        |Empty : Data.*)
    Definition TD := Time * Data.
    Variable Input : Stream TD.
    Variable Output : Stream TD.
\end{lstlisting}
%
In our framework, time is represented by real numbers. Benefit from the completeness of real number system, we can express and carry out the effective operation of a quantity at any precision request.
The continuity of the set of real numbers is sufficiently enough for our modeling approach. Also the
continuous time model is more appropriate since it is very expressive and closer to the nature of time in the real world. Thus, the time sequence consists of increasing and diverging time moments. For simplicity, here we take the natural numbers as the definition of data, which can be easily expanded according to different application domains. The Cartesian product of time and data defines a TD object.
We use the stream module in Coq to produce streams of TD objects.

@@

The simplest form of a synchronous channel is denoted by the Sync channel type. For a channel of the Sync type, a read operation on
its sink end succeeds only if there is a write operation pending on its source end. Thus, the time and data of a stream flowing
into the channel are exactly the same as the stream that flows out of the channel
\footnote{If we use $\alpha,\beta$ to denote the data streams that flow through the channel ends of a channel and $a,b$ to denote the time stream corresponding to the data streams, i.e., the $i$-th element $a(i)$ in $a$ denotes exactly the time moment of the occurrence of $\alpha(i)$, then we can easily obtain the specifications for different channels, as discussed in \cite{Sun12,SAA+12}. For example, a synchronous channel can be expressed as $\alpha=\beta \wedge \emph{a}=\emph{b}$.}.
The Sync channel can be defined as follows in the Coq system:
\begin{lstlisting}[language=coq]
    Definition Sync (Input Output:Stream TD) : Prop :=
    Teq Input Output /\ Deq Input Output.
\end{lstlisting}

The channel of type SyncDrain is a synchronous channel that allows pairs of write operations pending on its two ends to succeed simultaneously. All written data items are lost. Thus, the SyncDrain channel is used for synchronising two timed data streams on its two source ends. This channel type is an important basic synchronization building block for the construction of more complex connectors. The SyncDrain channel can be defined as follows:
\begin{lstlisting}[language=coq]
    Definition SyncDrain (Input Output:Stream TD) : Prop :=
    Teq Input Output.
\end{lstlisting}

The channel types FIFO and FIFO$n$ where $n$ is an integer greater than $0$ represent the
typical unbounded and bounded asynchronous FIFO channels. A write to a FIFO channel always succeeds, and a write to a FIFO$n$ channel succeeds only if the
number of data items in its buffer is less than its bounded capacity $n$. A read or take from a FIFO or FIFO$n$ channel suspends until the first data item
in the channel buffer can be obtained and then the operation succeeds.
%We use the $\overrightarrow{\emph{a}}$ to denote the tail of  a sequence \emph{a}.
For simplicity, we take the FIFO1 channel as an example. This channel type requires that the time when it consumes a data item through its source end is earlier than the time when the data item is delivered through its sink end. Besides, as the buffer has the capacity 1, time of the next data item that flows in should be later than the time when the data in the buffer is delivered. We use intersection of predicates in its definition as follows:
\begin{lstlisting}[language=coq]
    Definition FIFO1(Input Output:Stream TD) : Prop :=
    Tle Input Output /\ Tle Output (tl Input)
    /\ Deq Input Output.
\end{lstlisting}

For a FIFO1 channel whose buffer already contains a data element \emph{e}, the communication can be initiated only if the data element \emph{e} can be taken via the sink end. In this case, the data stream that flows out of the channel should get an extra element \emph{e} settled at the beginning of the stream. And time of the stream that flows into the channel should be earlier than time of the tail of the stream that flows out. But as the buffer contains the data element \emph{e}, new data can be written into the channel only after the element \emph{e} has been taken. Therefore, time of the stream that flows out is earlier than time of the stream that flows in. The channel can be represented as the \begin{lstlisting}[language=coq]
    Definition FIFO1e(Input Output:Stream TD)(e:Data) : Prop :=
    Tgt Input Output /\ Tle Input (tl Output)
    /\ PrR (hd Output) = e  /\ Deq Input (tl Output).
\end{lstlisting}

@@

A LossySync channel behaves the same as a Sync channel, except that a write operation on its source always succeeds immediately. If a compatible read or take operation is already pending on the sink of a LossySync channel, the written data item is transferred to the pending operation and both succeed. Otherwise, the write operation succeeds and the data item is lost. The LossySync channel can be defined as follows:
\begin{lstlisting}[language=coq]
   Parameter LossySync: Stream TD -> Stream TD -> Prop.
   Axiom LossySync_coind:
     forall Input Output: Stream TD,
     LossySync Input Output ->
       (
        (hd Output = hd Input  /\ LossySync (tl Input)(tl Output))
        \/
        LossySync(tl Input) Output
       ).
\end{lstlisting}

AsyncDrain is analogous to SyncDrain except that it guarantees that the pairs of write operations on the two channel ends never succeed simultaneously. Similarly it only has requirements on the time of the two streams on its opposite ends, but it requires that the times of the two streams are always different. The AsyncDrain channel can be defined as follows:
\begin{lstlisting}[language=coq]
    Parameter AsyncDrain: Stream TD -> Stream TD -> Prop.
    Axiom AsyncDrain_coind:
      forall Input1 Input2: Stream TD,
      AsyncDrain Input1 Input2 ->
        (~ PrL(hd Input1)  =  PrL (hd Input2) )
        /\
        ( (
          (PrL(hd Input1) < PrL (hd Input2)) /\
          AsyncDrain (tl Input1) Input2
          ) /\
          (
          (PrL(hd Input1) > PrL (hd Input2))  /\
          AsyncDrain Input1 (tl Input2)
          )
        ).
\end{lstlisting}

@@

The merge operator is more complicated. We consider merging two
channels \emph{AB} and \emph{CD}. When the merge operator acts on
these two channels, it leads to a choice of taking from the common
node that delivers a data item out of \emph{AB} or \emph{CD}.
%If both channels have data items available, the choice is
%non-deterministic\footnote{Choices are made by external context, so
%the definition here remains deterministic.}.
Similar to the definition of basic channels, we define merge as the intersection of two predicates and use recursive definition here:
%    Definition xor(a b: Prop) :=  (a \/ b) /\ ~(a /\ b).
\begin{lstlisting}[language=coq]
    Parameter merge:
    Stream TD -> Stream TD ->Stream TD -> Prop.
    Axiom merge_coind:
      forall s1 s2 s3:Stream TD,
      merge s1 s2 s3-> (
        ~ (PrL(hd s1) = PrL(hd s2)) /\
        (
          (PrL(hd s1) < PrL(hd s2)) ->
          ((hd s3 = hd s1)  /\ merge (tl s1) s2 (tl s3))
        ) /\ (
          (PrL(hd s1) > PrL(hd s2)) ->
          ((hd s3 = hd s2)  /\ merge s1 (tl s2) (tl s3))
        )
      ).
\end{lstlisting}

@@

In Coq, the theorem can be proved as follows:
\begin{lstlisting}[language=coq]
    Theorem test1: forall A B C,
      Sync A B /\ FIFO1 B C -> Tle A C.
    Proof.
        intros. destruct H. destruct H0.
        intro n. rewrite H. apply H0.
    Qed.
\end{lstlisting}

@@

We first introduce some lemmas to facilitate the proof.
\begin{lstlisting}[language=coq]
    Lemma transfer_eq : forall s1 s2 s3 : Stream TD,
    ((Teq s1 s2) /\ (Teq s2 s3)) -> (Teq s1 s3).
    Lemma transfer_eqtl : forall s1 s2 : Stream TD,
    (Teq s1 s2) -> (Teq tl s1)) (tl s2)).
    Lemma transfer_leeq : forall s1 s2 s3 : Stream TD,
    ((Tle s1 s2) /\ (Teq s2 s3)) -> (Tle s1 s3).
    Lemma transfer_hdle : forall s1 s2 : Stream TD,
    (Tle s2 s1) -> (PrL (hd s1) > PrL (hd s2)).
\end{lstlisting}

@@

\ionode{(io1)}{(3,1.2)}{node[left]{A}}
\ionode{(io2)}{(3,0)}{node[left]{B}}
\ionode{(io3)}{(6.5,1.2)}{node[right]{C}}
%\ionode{(io4)}{(5,0)}{node[right]{C2}}
\syncdrain{(io1)}{(io2)}{}
\fifoe{(io1)}{(io3)}{}
\sync{(io2)}{(io3)}{}
\draw (4.7, -0.6) node[draw=none] {(b) After merge};
\end{tikzpicture}
\caption{Alternator}
\label{fig:alternator}
\end{figure}

@@

\begin{lstlisting}[language=coq]
    Section Alt.
    Hypothesis D1: SyncDrain A B.
    Hypothesis D2: FIFO1 A C1.
    Hypothesis D3: Sync B C2.
    Theorem subtest:
       (Tle C2 C1)  /\  (Tle C1 (tl C2)).
\end{lstlisting}

@@

An additional hypothesis is needed for the proof of alternator which merges \emph{C1} and \emph{C2} into a common node \emph{C}. Based on the three
hypotheses for channels and the additional hypothesis, the theorem of alternator is presented as the following proposition which needs to be proved:
\begin{lstlisting}[language=coq]
    Hypothesis D4: merge C1 C2 C.
    Theorem test:
    hd(C) = hd(C2) /\ merge C1 (tl C2) (tl C).
    Proof.
      destruct subtest.  (* ... *)
\end{lstlisting}
Here we only present the first step which shows how a proven theorem can be applied in another proof and omit the full details because of the page limitation.
And it greatly simplifies the process of proving the property of alternator.
\end{example}

@@

\begin{lstlisting}[language=coq]
    Theorem refinement : forall A C D,
    (exists B, (FIFO1 A B) /\ (Sync B C) /\ (Sync B D)) ->
    (exists E, (Sync A E) /\ (FIFO1 E C) /\ (FIFO1 E D)).
\end{lstlisting}

@@

Equivalence relationship between the two connectors can be formalized as:
\begin{lstlisting}[language=coq]
    Theorem equivalence: forall A B C,
      (exists E F G,
        (Sync A E) /\ (FIFO1 E F) /\ (Sync F B) /\
        (FIFO1 E G) /\ (Sync G C) /\ (SyncDrain F G)
      ) <->
      (exists D,
        (FIFO1 A D) /\ (Sync D B) /\ (Sync D C)
      ).
\end{lstlisting}
The proof of this theorem has two steps. Firstly, we prove that the new connector \emph{R} is a refinement of connector \emph{Q}.
We hope to find an appropriate \emph{$D$} that satisfies
\[
FIFO1(A,D)\wedge Sync(D,B) \wedge Sync(D,C)
\]
Similar to Example \ref{ex:refine}, we first assert $D=F$, which leads to
\[
FIFO1(A,F) \land Sync(F,B) \wedge Sync(F,C)
\]
From Lemma \ref{lemma:eq}, we have $Sync(A,E)$, or $A=E$. Therefore, \emph{FIFO1(E,F)} can be replaced by \emph{FIFO1(A,F)}. By adopting \emph{FIFO1(E,F)} and \emph{FIFO1(E,G)}, we can prove that the data sequences at \emph{F} and \emph{G} are equal. Similarly, data sequences at $C$, $G$ and $F$ are also equal, wrt. $Sync(G,C)$.

@@

By means of function \emph{Conjunction}, basic channels are defined concisely and neatly. For example, the specific behavior constraints for \emph{Sync} channel are classified into two types: data constraints and time constraints. The description and the definition of \emph{Sync} in Coq have already clearly expressed the behavior pattern it needs to follow: each item in data and time streams of output specified as ``node[1]'' is equivalent to the timed data items of input specified as ``node[0]'', which are exactly data and time constraints. As they are both requirements that we need to satisfy no matter data or time is related, each constraint in the list will all be merged finally.
\lstset{language=Python}
\begin{lstlisting}[frame=single]
def Sync(nodes, bound):
   assert len(nodes) == 2
   constraints = []
   for i in range(bound):
     constraints += [ nodes[0]['data'][i] == nodes[1]['data'][i] ]
     constraints += [ nodes[0]['time'][i] == nodes[1]['time'][i] ]
   return Merge(constraints)
\end{lstlisting}

@@

c2 = Connector()
c2.connect('Fifo1', 'A', 'N')
c2.connect('Sync', 'N', 'B')
c2.connect('Sync', 'N', 'C')
\end{lstlisting}
Notice that the \emph{replicate} operator is applied twice in the construction process. Assume the bound is ten and then carry out the function \emph{c2.isRefinementOf(c1, 10)}. Finally, we can get the result \emph{True} which means that \emph{Connector Q} in Fig.\ref{refine} is indeed a refinement of \emph{Connector P}. We can further extend the bound and the time complexity increases linearly as no recursively defined channels are involved.
\end{example}
\begin{example}
This example checks the equivalence relation between connectors in Fig.\ref{refine1}. Construct the two connectors first similar to the example \ref{ex:equivalence1} and invoke the function \emph{isRefinementOf} in two directions. 
When giving the bound ten, the results of two directions that we get are both \emph{True}. With the yielded result and the soundness theorem, we can conclude that the two connectors are in equivalence relation within the bound ten.
\end{example}
